---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---


## Loading and preprocessing the data
Loading activity.csv file into data frame .We will need to convert date column to proper date format. We cann use as.date function for this.

```{r loaddata}

rep_dta<-read.csv("activity.csv")
rep_dta$date<-as.Date(as.character(rep_dta$date))
```


## What is mean total number of steps taken per day?

Using tapply Funciton to calculate Aberage steps per day. Then we can use host function to show histogram.

```{r calstepsperday}
steps_per_day<-tapply(rep_dta$steps,rep_dta$date,FUN=sum,na.rm=TRUE)

hist(steps_per_day,ylab="Day",col="green",breaks = 20)

```

Mean Total Number of steps day 
```{r meanstesperday}
steps_per_day_mean<-tapply(rep_dta$steps,rep_dta$date,FUN=mean,na.rm=TRUE)

print(steps_per_day_mean)


```
Median Total Number of steps day 
```{r medianstepsperday}
steps_per_day_median<-tapply(rep_dta$steps,rep_dta$date,FUN=median,na.rm=TRUE)

print(steps_per_day_median)


```

## What is the average daily activity pattern?

We can use tapply function to calculate Average no of steps for each 5 min interval .We stored thsis data into new variable interval_dta. We can use the same to plot .
```{r dailyactivitypattern}

interval_dta<-tapply(rep_dta$steps,rep_dta$interval,FUN=mean,na.rm=T)
plot(interval_dta,xlab="5 min interval", ylab="Avg steps taken ",main="Avg Steps taken by each 5 min interval")

```
## Imputing missing values

Total Number of Missing values in dataset
```{r missingvalues}

sum(is.na(rep_dta))


```

Creating new dataset with  missing values filled 

```{r cleardataset}
new_dta<-rep_dta
new_dta[is.na(new_dta$steps),1]<-mean(new_dta$steps,na.rm = TRUE)
steps_per_day_new<-tapply(new_dta$steps,new_dta$date,FUN=sum)
```
Histgram with new dataset

```{r histnew}
hist(steps_per_day_new,ylab="Day",col="green",breaks = 20)

```


Mean Total Number of steps day for new dataset

```{r meanstesperdaynew}
steps_per_day_mean_new<-tapply(new_dta$steps,new_dta$date,FUN=mean,na.rm=TRUE)

print(steps_per_day_mean_new)


```
Median Total Number of steps day for new dataset
```{r medianstepsperdaynew}
steps_per_day_median_new<-tapply(new_dta$steps,new_dta$date,FUN=median,na.rm=TRUE)

print(steps_per_day_median_new)


```

*Replacing Missing values did affect the meaan and median of datasets. With replacing NA  with mean ,count of zero values reduced .data is now normall distributed. 



## Are there differences in activity patterns between weekdays and weekends?
We can seperate data with the help of weekdays function . We have added one more column as Day to dataset to show whether it is weekday or weekend.

We can plot two line plots seperatly and see comparison between two categories.

```{r compareweekdayweekend}
new_dta$day<-weekdays(new_dta$date)
new_dta[new_dta$day=="Monday"|new_dta$day=="Tuesday"|new_dta$day=="Wednesday"|new_dta$day=="Thursday"|new_dta$day=="Friday",4]<-"Weekday"
new_dta[new_dta$day=="Sunday"|new_dta$day=="Saturday",4]<-"Weekend"
wkd<-new_dta[new_dta$day=="Weekday",]
wkn<-new_dta[new_dta$day=="Weekend",]

interval_dta_weekday<-tapply(wkd$steps,wkd$interval,FUN=mean,na.rm=T)


interval_dta_weekend<-tapply(wkn$steps,wkn$interval,FUN=mean,na.rm=T)
par(mfrow=c(2,1),mar=c(2,4,2,2))
plot(interval_dta_weekend,type="l",main="Weekend",xlab="Interval",ylab="Number of Steps")
plot(interval_dta_weekday,type="l",main="Weekday",xlab="Interval",ylab="Number of Steps")

```



